---
title: 'Long term trends: Ovigerous female sea lice'
author: "Bela Klimesova"
date: "2025-05-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Mean sea lice dataset (SL_Mean)

```{r}
SL_Mean<-read.table("MeanDataset.csv",
	header = TRUE,sep = ',',fill = TRUE,dec = ".",na.strings = "NA")
```

```{r}
library(ggplot2)
library(glmmTMB)
suppressMessages(library(lmtest))
library(performance)
library(gridExtra)
library(grid)
```

# Data adjustment before analysis

The outliers were identified: sites with less than 120 inspections, fish that stayed more than 24 months in the seawater, and outliers (total mobile sea lice mean values higher than 90).

```{r, include=FALSE}
# Histogram of high sea lice counts
hist(SL_Mean$MeanLepEggs,ylim=c(0,100))

# Number of observation from fish more than 24 months in the seawater
nrow(SL_Mean[SL_Mean$TimeSea==30,])
```

```{r}
# Excluding fish stock transported to sea water in 2023
SL_Mean1<-SL_Mean[!(SL_Mean$New_Class %in% c("2023 S1/2","2023 S1")),]

# Sites with less than 120 inspections
Insp<-aggregate(SL_Mean1$SplDate,list(SL_Mean1$SiteName),function(x){length(unique(x))})
Insp_Sel<-Insp[Insp$x<120,]$Group.1

# Fish more than 24 months in the sea water, group '30' includes 25-30 in seawater
SL_Mean2<-SL_Mean1[SL_Mean1$TimeSea<30,]
SL_Mean3<-SL_Mean2[!(SL_Mean2$SiteName %in% Insp_Sel),]

# Outliers
SL_Pen<-SL_Mean3[SL_Mean3$MeanLepEggs<40,]
```

Comparison of raw and cleaned datasets

```{r}
# Number of counts per fish
nrow(SL_Mean)
nrow(SL_Pen)

# Number of salmon farms
length(unique(SL_Mean$SiteName))
length(unique(SL_Pen$SiteName))
```

```{r}
# Number of mean sea lice values per pen in the dataset
nrow(SL_Pen)
mean(SL_Pen$MeanLepTotal);sd(SL_Pen$MeanLepTotal)

# Proportion of Total mobile and ovigerous female sea lice
sum(SL_Pen$MeanLepEggs)/sum(SL_Pen$MeanLepTotal)
```

The format of variables in the dataset (SL_Pen), sea lice values recorded as a mean value per fish per farm, was adjusted. 

```{r}
SL_Pen$Year_f<-as.factor(SL_Pen$Year)
SL_Pen$TimeSea_f<-as.factor(SL_Pen$TimeSea)
SL_Pen$Year_n<-as.numeric(SL_Pen$Year_f)/10
SL_Pen$NewMonth<-factor(SL_Pen$NewMonth,levels=c('Jan/Dec','Feb','Mar1','Mar2',
	'Apr1','Apr2','May1','May2','Jun','Jul','Aug','Sep','Oct','Nov'))
```

A new dataset (SL_Farm), sea lice values recorded as a mean value per fish per farm, was created. 

```{r}
SL_Farm<-aggregate(SL_Pen$MeanLepEggs,list(SL_Pen$Year,SL_Pen$Year_f,
              SL_Pen$SiteName,SL_Pen$NewMonth,SL_Pen$Month,SL_Pen$Class,
              SL_Pen$TimeSea_f,SL_Pen$Latitude,SL_Pen$Longitude,SL_Pen$SplDate),mean)
colnames(SL_Farm)<-c('Year','Year_f','SiteName','NewMonth','Month','Class',
                     'TimeSea','Latitude','Longitude','Date','Lep')
```

The format of variables in the dataset (SL_Farm) was adjusted.

```{r}
SL_Farm$TimeSea<-as.factor(SL_Farm$TimeSea)

# Number of mean values per farm in the dataset
nrow(SL_Farm)
```

# Year as categorical predictor

The temporal autocorrelation for model at the pen (mean value of sea lice per fish per pen) and at the farm (mean value of sea lice per fish per farm) level were compared.

## Pen level

The SL_Pen dataset (mean value per fish per pen) was used in a GLM that included all relevant predictors.

```{r}
m5<-glmmTMB(MeanLepEggs~Year_f+NewMonth+TimeSea_f+Class+CageType+scale(Latitude)+
              scale(Longitude)+(1|SiteName),ziformula=~Year_f+NewMonth+
              TimeSea_f+Class+CageType+scale(Latitude)+scale(Longitude),
              family=ziGamma(link="log"),data=SL_Pen)
```

The temporal autocorrelation was tested.

```{r}
dwtest(m5)
bgtest(m5)
```

## Farm level

The SL_Farm dataset (mean value per fish per farm) was used in a GLM that included all relevant predictors.

```{r}
m6<-glmmTMB(Lep~Year_f+NewMonth+TimeSea+Class+scale(Latitude)+
            scale(Longitude)+(1|SiteName),ziformula=~
            Year_f+NewMonth+TimeSea+Class+scale(Latitude)+scale(Longitude),
            family=ziGamma(link="log"),data=SL_Farm)
```

The model was then tested for temporal autocorrelation and the results were compared to the model (m5) at the pen level.

```{r}
dwtest(m6)
bgtest(m6)
```

Lower temporal autocorrelation was detected for the model at the pen level. Based on statistical diagnostics, data availability, and the research question the pen level data (SL_Pen) were used for further analysis.

The model was checked for multicollinearity. 

```{r}
check_collinearity(m5)
```

The predictors were selected using a backwards step-wise approach.

```{r}
m5.1<-glmmTMB(MeanLepEggs~NewMonth+TimeSea_f+Class+CageType+scale(Latitude)+
              scale(Longitude)+(1|SiteName),ziformula=~NewMonth+
              TimeSea_f+Class+CageType+scale(Latitude)+scale(Longitude),
              family=ziGamma(link="log"),data=SL_Pen)
m5.2<-glmmTMB(MeanLepEggs~Year_f+TimeSea_f+Class+CageType+scale(Latitude)+
              scale(Longitude)+(1|SiteName),ziformula=~Year_f+
              TimeSea_f+Class+CageType+scale(Latitude)+scale(Longitude),
              family=ziGamma(link="log"),data=SL_Pen)
m5.3<-glmmTMB(MeanLepEggs~Year_f+NewMonth+Class+CageType+scale(Latitude)+
              scale(Longitude)+(1|SiteName),ziformula=~Year_f+NewMonth+
              Class+CageType+scale(Latitude)+scale(Longitude),
              family=ziGamma(link="log"),data=SL_Pen)
m5.4<-glmmTMB(MeanLepEggs~Year_f+NewMonth+TimeSea_f+CageType+scale(Latitude)+
              scale(Longitude)+(1|SiteName),ziformula=~Year_f+NewMonth+
              TimeSea_f+CageType+scale(Latitude)+scale(Longitude),
              family=ziGamma(link="log"),data=SL_Pen)
m5.5<-glmmTMB(MeanLepEggs~Year_f+NewMonth+TimeSea_f+Class+scale(Latitude)+
              scale(Longitude)+(1|SiteName),ziformula=~Year_f+NewMonth+
              TimeSea_f+Class+scale(Latitude)+scale(Longitude),
              family=ziGamma(link="log"),data=SL_Pen)
m5.6<-glmmTMB(MeanLepEggs~Year_f+NewMonth+TimeSea_f+Class+CageType+
              scale(Longitude)+(1|SiteName),ziformula=~Year_f+NewMonth+
              TimeSea_f+Class+CageType+scale(Longitude),
              family=ziGamma(link="log"),data=SL_Pen)
m5.7<-glmmTMB(MeanLepEggs~Year_f+NewMonth+TimeSea_f+Class+CageType+scale(Latitude)+
              (1|SiteName),ziformula=~Year_f+NewMonth+
              TimeSea_f+Class+CageType+scale(Latitude),
              family=ziGamma(link="log"),data=SL_Pen)

AIC(m5,m5.1,m5.2,m5.3,m5.4,m5.5,m5.6,m5.7)
BIC(m5,m5.1,m5.2,m5.3,m5.4,m5.5,m5.6,m5.7)
```

Best fit was determined for the full model m5. 
The final model was then tested for overdispersion and the explained variance was determined. 

```{r}
# Overdispersion
E2 <- resid(m5, type = "pearson")
N  <- nrow(SL_Pen)
p  <- length(coef(m5))   
sum(E2^2) / (N - p)

# Explained variance
r2_nakagawa(m5)
```

The results of the final model were checked.

```{r}
summary(m5)
car::Anova(m5)
```

## Visualization of results

The sea lice values for all levels of predictors with the confidential interval were predicted.

```{r}
newdata2 <- data.frame(
	Class=factor(rep(rep(rep(rep(rep(rep(1:2),13),31),14),4),2),levels=1:2,
	             labels=levels(as.factor(SL_Pen$Class))),
	Year_f=factor(rep(rep(rep(rep(1:31,each=26),14),4),2),levels=1:31,
	              labels=levels(SL_Pen$Year_f)),
	NewMonth=factor(rep(rep(rep(1:14,each=806),4),2),levels=1:14,
	                labels=levels(SL_Pen$NewMonth)),
	TimeSea_f=factor(rep(rep(1:4,each=11284),2),levels = 1:4,
	                 labels =levels(SL_Pen$TimeSea_f)),
	CageType=factor(rep(1:2,each=45136),levels = 1:2,
	                 labels =levels(as.factor(SL_Pen$CageType))),
	Latitude=mean(SL_Pen$Latitude),
	Longitude=mean(SL_Pen$Longitude))

newdata2 <- cbind(newdata2, predict(m5, newdata2, type = "link", se.fit=TRUE,re.form=NA))

newdata2 <- within(newdata2, {
  Abun<- exp(fit)
  LL <- exp(fit - 1.96 * se.fit)
  UL <- exp(fit + 1.96 * se.fit)
})

```

Mean values of sea lice year were calculated from predicted sea lice counts.

```{r}
A<-aggregate(newdata2$Abun,list(newdata2$Year_f),mean)
B<-aggregate(newdata2$LL,list(newdata2$Year_f),mean)
C<-aggregate(newdata2$UL,list(newdata2$Year_f),mean)
colnames(A)<-c('Year','Abun')
colnames(B)<-c('Year','LL')
colnames(C)<-c('Year','UL')
AA<-merge(A,B, by="Year")
BB<-merge(AA,C, by="Year")
```

The final data were visualized together with the raw data for every year at the level of farm

```{r}
BB$Year<-as.numeric(1993:2023)

# Predicted sea lice counts
A<-ggplot(BB, aes(Year, Abun)) +
	geom_ribbon(aes(ymin = LL, ymax = UL), colour = 'grey80', fill = "grey80") +
	geom_point() +
	geom_line(linewidth = 1.1) +
	ylab('[annual mean per fish per pen]')+
	theme_minimal(base_size = 10) +
	theme(legend.position = "none",
		panel.grid.minor = element_blank()) +
	scale_x_continuous(name = "Year",limits = c(1992.5, 2023.5),
		breaks = seq(1993, 2023, 5)) +
	coord_cartesian(xlim = c(1993, 2023))

# Observed sea lice counts
SL_Pen$Year <- as.numeric(as.character(SL_Pen$Year_f))
B<-ggplot(SL_Pen, aes(x = Year, y = MeanLepEggs, group = Year)) +
	geom_boxplot() +
	scale_y_sqrt() +
	ylab('[sqrt mean per fish per pen]')+
	theme_minimal(base_size = 10) +
	theme(legend.position = "none",
		axis.title.x = element_blank(),
		axis.text.x = element_blank(),
		panel.grid.minor = element_blank()) +
	scale_x_continuous(limits = c(1992.5, 2023.5),
		breaks = seq(1993, 2023, 5)) +
	coord_cartesian(xlim = c(1993, 2023))

grid.arrange(B,A,ncol = 1,
  heights = c(1, 1.2),left = textGrob("Ovigerous female sea lice abundance",
    gp = gpar(fontsize = 15),rot = 90))

# Adding labels A and B
grid.text("a", x = unit(0.10, "npc"), y = unit(.968, "npc"),
          gp = gpar(fontsize = 15, col = "black"))
grid.text("b", x = unit(0.10, "npc"), y = unit(.51, "npc"),
          gp = gpar(fontsize = 15, col = "black"))
```

The predicted values were used to calculate mean values per month to visualize the seasonal changes in sea lice abundance.

```{r}
newdata2$NewMonth_n<-substr(as.character(newdata2$NewMonth),1,3)
newdata2[newdata2$NewMonth_n=='Jan',]$NewMonth_n<-'Jan/Dec'

A<-aggregate(newdata2$Abun,list(newdata2$NewMonth_n),mean)
B<-aggregate(newdata2$LL,list(newdata2$NewMonth_n),mean)
C<-aggregate(newdata2$UL,list(newdata2$NewMonth_n),mean)
colnames(A)<-c('Month','Abun')
colnames(B)<-c('Month','LL')
colnames(C)<-c('Month','UL')
AA<-merge(A,B, by="Month")
BB<-merge(AA,C, by="Month")

# Common x-axis labels
BB$Month<-factor(BB$Month,levels=c("Jan/Dec","Feb","Mar",
	"Apr","May","Jun","Jul","Aug","Sep","Oct","Nov"))
BLUE <- data.frame(X=c('Mar','Apr','May'), Y = 3)

# Predicted sea lice counts
ggplot(data=BB,aes(Month,Abun,group=1))+
	geom_area(aes(x=X,y=Y),data=BLUE,alpha=0.3,fill="grey80") +
	geom_ribbon(aes(ymin = LL, ymax=UL),colour ='grey80',fill="grey80") +
	geom_point() +
	geom_line(linewidth = 1.1)+
	theme_minimal(base_size =15) + 
	ylab('Ovigerous female sea lice abundance \n [monthly mean per fish per pen]')+
	xlab('Month')+
	ylim(0,3)
```

# Year as numeric

## Pen level

The dataset at the pen level was used to fit a model including all predictors. 

```{r}
m7<-glmmTMB(MeanLepEggs~Year+NewMonth+TimeSea_f+Class+CageType+scale(Latitude)+
              scale(Longitude)+(1|SiteName),ziformula=~Year+NewMonth+
              TimeSea_f+Class+CageType+scale(Latitude)+scale(Longitude),
            family=ziGamma(link="log"),data=SL_Pen)
```

The temporal autocorrelation was determined.

```{r}
dwtest(m7)
bgtest(m7)
```

## Farm level

The dataset at the farm level was used to fit a model including all predictors.

```{r}
m8<-glmmTMB(Lep~Year+NewMonth+TimeSea+Class+scale(Latitude)+
              scale(Longitude)+(1|SiteName),ziformula=~Year+
              NewMonth+TimeSea+Class+scale(Latitude)+scale(Longitude),
              family=ziGamma(link="log"),data=SL_Farm)
```

The temporal autocorrelation was determined and compared to the model at the pen level.

```{r}
dwtest(m8)
bgtest(m8)
```

Lower temporal autocorrelation was detected for the model at the pen level. Based on statistical diagnostics, data availability, and the research question the pen level data (SL_Pen) were used for further analysis.

```{r}
check_collinearity(m7)
```

The predictors were selected using a backwards step-wise approach.

```{r}
m7.1<-glmmTMB(MeanLepEggs~NewMonth+TimeSea_f+Class+CageType+scale(Latitude)+
              scale(Longitude)+(1|SiteName),ziformula=~NewMonth+
              TimeSea_f+Class+CageType+scale(Latitude)+scale(Longitude),
            family=ziGamma(link="log"),data=SL_Pen)
m7.2<-glmmTMB(MeanLepEggs~Year+TimeSea_f+Class+CageType+scale(Latitude)+
              scale(Longitude)+(1|SiteName),ziformula=~Year+
              TimeSea_f+Class+CageType+scale(Latitude)+scale(Longitude),
            family=ziGamma(link="log"),data=SL_Pen)
m7.3<-glmmTMB(MeanLepEggs~Year+NewMonth+Class+CageType+scale(Latitude)+
              scale(Longitude)+(1|SiteName),ziformula=~Year+NewMonth+
              Class+CageType+scale(Latitude)+scale(Longitude),
            family=ziGamma(link="log"),data=SL_Pen)
m7.4<-glmmTMB(MeanLepEggs~Year+NewMonth+TimeSea_f+CageType+scale(Latitude)+
              scale(Longitude)+(1|SiteName),ziformula=~Year+NewMonth+
              TimeSea_f+CageType+scale(Latitude)+scale(Longitude),
            family=ziGamma(link="log"),data=SL_Pen)
m7.5<-glmmTMB(MeanLepEggs~Year+NewMonth+TimeSea_f+Class+scale(Latitude)+
              scale(Longitude)+(1|SiteName),ziformula=~Year+NewMonth+
              TimeSea_f+Class+scale(Latitude)+scale(Longitude),
            family=ziGamma(link="log"),data=SL_Pen)
m7.6<-glmmTMB(MeanLepEggs~Year+NewMonth+TimeSea_f+Class+CageType+
              scale(Longitude)+(1|SiteName),ziformula=~Year+NewMonth+
              TimeSea_f+Class+CageType+scale(Longitude),
            family=ziGamma(link="log"),data=SL_Pen)
m7.7<-glmmTMB(MeanLepEggs~Year+NewMonth+TimeSea_f+Class+CageType+scale(Latitude)+
              (1|SiteName),ziformula=~Year+NewMonth+
              TimeSea_f+Class+CageType+scale(Latitude),
            family=ziGamma(link="log"),data=SL_Pen)

AIC(m7,m7.1,m7.2,m7.3,m7.4,m7.5,m7.6,m7.7)
BIC(m7,m7.1,m7.2,m7.3,m7.4,m7.5,m7.6,m7.7)
```

Best fit was determined for the full model m7. 
The final model was then tested for overdispersion and the explained variance was determined. 

```{r}
# Overdispersion
E2 <- resid(m7, type = "pearson")
N  <- nrow(SL_Pen)
p  <- length(coef(m7))   
sum(E2^2) / (N - p)

# Explained variance
r2_nakagawa(m7)
```

The results of the final model were checked.

```{r}
summary(m7)
car::Anova(m7)
```
